### 2010-12-24 Documentation/atropos/atropos.tex
ENH: Bweb discussion and associated references
git-svn-id: https://advants.svn.sourceforge.net/svnroot/advants@710 fd2da0a4-d968-48df-9cb5-d8c496287cbf

  initialization is competitive with but inferior to K-means.}

@@

\subsection{Performance on Single-Subject Brainweb Data}
How does Atropos performance compare to other methods that have been
evaluated on the basic brainweb data?   {\bf {\textcolor{red}{ FIXME}}}.

\subsection{Performance on Twenty-Subject Brainweb Data}
How does Atropos performance compare to other methods that have been
evaluated on the brainweb twenty data?   {\bf {\textcolor{red}{ FIXME}}}.

@@

\subsection{Other Thoughts} {\bf {\textcolor{red}{ FIXME}}}
A more general advantage which extends beyond the scope of the experimental 
evaluation section of this paper is the flexibility of the Atropos algorithm.   This includes 
not only $n$-tissue segmentation and dense volumetric cortical parcellation,
as reported in this work, but Atropos is also used in conjunction with our ANTs registration tools for robust
brain extraction which has reported good performance in comparison with other
popular, publicly available brain extraction tools \citep{Avants2010a}.  An additional successful
application area is that of ventilation-based segmentation of hyperpolarized helium-3 MRI
\citep{Tustison2010a}.

\section{Conclusion}
The Atropos algorithm is freely available to the public.  We release
this code not only to make it available to clinical researchers but
with the hope that other researchers in segmentation will provide
feedback about the implementation decisions that we made.  EM
segmentation is non-trivial and there are numerous design alternatives
available not only in the models selected but also in the ICM coding
strategy, alternatives to ICM and the specific method in which prior
and likelihood are combined.  Due to the flexibility of Atropos, we
also hope that some of its capabilities, though not evaluated here,
are explored by the segmentation or clinical community.

>>>>

  initialization is competitive with but inferior to K-means.  The
  legend is in the same position in each graph, allowing a visual
  comparison of the results.  As one may see, the N4 assisted overlap
  values are consistent across bias field/RF inhomogeneity.}

@@

\subsection{Performance on Brainweb Data}
Atropos results are competitive with the state of the art.  For
instance, \cite{Ashburner2005} (SPM5) evaluated on 0\% RF (bias field) 3\%
noise braineweb single subject data finding 0.932 (GM) and .961 (WM)
Dice. Results on 40\% RF were .934 (GM) and .961 (WM). SPM5 exhibits
insensitivity to bias similar to our own best results on the 40\%RF, 3
\% noise case (MRF-$\beta$=0.2, Kmeans $+$ N4) with Dice for GM is
0.950737 and for WM is 0.962586.  \cite{Nakamura2009} gave GM Dice
results (brainweb single 3\% noise) of 0.962 (0\%RF), 0.964(20\%RF)
and .956(40\%RF) which are slightly higher than either SPM5 or Atropos
results.  However, \cite{Nakamura2009} do not report WM or CSF numbers
for comparison.  Topology preserving approaches also perform well.
\cite{Shiee2010} achieved Dice results for 3\% noise 20\% RF brainweb
single subject with 0.927, 0.912 and 0.900 for WM, GM and CSF,
respectively.  These are excellent numbers given the additional
constraint applied to the segmentation.  
\cite{Bazin2007c} proposed the TOADS method and, estimating from the
paper's graph, showed that the average Dice accuracy for 3\% noise various RF was
.95-.96 for WM, .93-.95 for GM and .92-.94 for CSF.
Perhaps the most recent balanced evaluation was performed in
\citep{Klauschen2009}, which reports confusion matrix numbers, rather
than Dice overlap.  Because the absolute true number of GM and WM
voxels for brainweb are known, we can convert the confusion matrix to
Dice.  In that case, the SPM5 Dice for brainweb single GM and WM is .8847
and .9085, respectively, while FreeSurfer and FSL's accuracy is
lower. The best GM result for 20 subjects is obtained by SPM5: 0.9295;
best WM Dice is from FSL: 0.9499.  We note that \cite{Klauschen2009}
used a comprehensive evaluation where quality of brain extraction also
contributed to the outcome.  Thus, the results must be interpreted
slightly differently than those from other papers.

@@

\subsection{Clinically-Related Evaluation}
While specifying performance on brainweb is highly
valuable, clinical validation is a second important aspect of
segmentation evaluation.  For instance, 
\citep{Freeborough1997,Westlye2009,Sanchez-Benavides2010,Chou2009,Bresser2011}
are only a few of the papers that evaluate segmentation performance with
respect to a known neurobiological outcome measure.  Atropos is
currently used in clinical studies and a number of clinically-focused,
application-specific evaluations are ongoing and will constitute
future work.  One early example of a clinically-focused Atropos
neuroimaging application is in \citep{Avants2010c}.  An additional successful
application area is that of ventilation-based segmentation of hyperpolarized helium-3 MRI
\citep{Tustison2010a} which also used the open-source  Glamorous Glue
algorithm to impose topology constraints.  Thus, future work may
incorporate topology more closely into the Atropos methodology.

\section{Conclusion}
A more general advantage which extends beyond the scope of the experimental 
evaluation section of this paper is the flexibility of the Atropos algorithm.   This includes 
not only $n$-tissue segmentation and dense volumetric cortical parcellation,
as reported in this work, but Atropos is also used in conjunction with our ANTs registration tools for robust
brain extraction which has reported good performance in comparison with other
popular, publicly available brain extraction tools \citep{Avants2010a}.

The Atropos algorithm is freely available to the public.  We release
this code not only to make it available to clinical researchers but
with the hope that other researchers in segmentation will provide
feedback about the implementation decisions that we made.  EM
segmentation is non-trivial and there are numerous design alternatives
available not only in the models selected but also in the ICM coding
strategy, alternatives to ICM and the specific method in which prior
and likelihood are combined.  Due to the flexibility of Atropos, we
also hope that some of its capabilities, though not evaluated here,
are explored by the segmentation or clinical community.
